# Causal Graphs {#chapter-2}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      results='hide', 
                      cache=FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      error = TRUE,
                      cache.comments = FALSE,
                      cache.lazy = FALSE,
                      fig.path = "bookdown-dag_files/figure-html/ch2-fig-",
                      cache.path = "bookdown-dag_cache/html/ch2-")

```

## The Fork

We first simulate some data from the simple 'Fork' DAG.

```{r}
set.seed(1747)
n <- 1e4
bZ <- 1
Z <- rnorm(n, 0, 1)
X <- Z*bZ + rnorm(n, 0, 1)
Y <- Z*bZ + rnorm(n, 0, 1)
```

We then write a function for fitting and plotting our models that we can re-use for the 'Pipe' and 'Collider' scenarios. This function depends on the `ggplot2` [@tidyverse] and `patchwork` [@patchwork] packages.

```{r}
library(ggplot2)
library(patchwork)

plot_scat <- function(data, title){

# Y ~ X
p1 <- ggplot(data, aes(x=X, y=Y)) + 
  geom_point(alpha = 0.05, size = .1) + 
  geom_smooth(method='lm', color = "blue", linewidth = 0.5) + 
  theme_classic() + 
  labs(title = title, subtitle = "Y ~ X")

# Y ~ X + Z
model <- lm(Y ~ X + Z, data = data)
new_data <- transform(data,
                      Z = 0)
predictions <- predict(model, newdata = new_data, interval = "confidence")

p2 <- ggplot(data, aes(x = X, y = Y)) +
  geom_point(alpha = 0.05, size = .1) + 
  geom_line(data = new_data, aes(y = predictions[, "fit"]), linewidth = 0.5, color = "blue") +
  geom_ribbon(data = new_data, aes(ymin = predictions[, "lwr"], ymax = predictions[, "upr"]), alpha = 0.2) +
  theme_classic() + 
  labs(subtitle = "Y ~ X + Z")

return(p1 + p2)
}
```

We then apply the function to the simulated data.

```{r, fig.width=5, fig.height=3}
plot_scat(data = data.frame(Y=Y, X=X, Z=Z), title = "The Fork")
```

## The Pipe

Similarly to above, we simulate data from the 'Pipe' DAG...

```{r}
set.seed(1747)
n <- 1e4
bZ <- 1
bX <- 1 
X <- rnorm(n, 0, 1)
Z <- X*bX + rnorm(n, 0, 1)
Y <- Z*bZ + rnorm(n, 0, 1)
```

... and then apply our custom fitting and plotting function.

```{r, fig.width=5, fig.height=3}
plot_scat(data = data.frame(Y=Y, X=X, Z=Z), title = "The Pipe")
```

## The Collider

Exactly the same approach as above.

```{r, fig.width=5, fig.height=3}
set.seed(1747)
n <- 1e4
bX <- 1
bY <- 1 
X <- rnorm(n, 0, 1)
Y <- rnorm(n, 0, 1)
Z <- X*bX + Y*bY + rnorm(n, 0, 1)

plot_scat(data = data.frame(Y=Y, X=X, Z=Z), title = "The Collider")
```

## Post-treatment bias

Again, following the same approach, we simulate data from the post-treatment DAG.

```{r}
set.seed(1747)
n <- 1e4
bZ <- 1
bX <- 0.5
bY <- 1
Z <- rnorm(n, 0, 1)
X <- Z*bZ + rnorm(n, 0, 1)
Y <- Z*bZ + X*bX + rnorm(n, 0, 1)
P <- Y*bY + rnorm(n, 0, 1)
```

Next, we fit two models: One that adjusts for the post-treatment variable *P*...

```{r, results='markup'}
glm(Y ~ X + Z + P, data = data.frame(Y=Y, X=X, Z=Z, P=P))
```

... and another that doesn't. 

```{r, results='markup'}
glm(Y ~ X + Z, data = data.frame(Y=Y, X=X, Z=Z))
```

We see that only the latter model picks up the correct estimate for *X*, which was 0.5 in this case.

## Session info

```{r, results='markup'}
sessionInfo()
```

