# Adventures in G-methods {#chapter-4}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      results='hide', 
                      cache=FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      error = TRUE,
                      cache.comments = FALSE,
                      cache.lazy = FALSE,
                      fig.path = "bookdown-dag_files/figure-html/ch4-fig-",
                      cache.path = "bookdown-dag_cache/html/ch4-")

```

## Doubly robust estimation

For demonstrating a 'doubly robust' estimator that combines IPW and g-computation, we use the `nhefs` data from the `causaldata` package [@causaldata]. This data come from the National Health and Nutrition Examination Survey Data I Epidemiologic Follow-up Study. 

```{r}
library(causaldata)
d <- nhefs
```

We first calculate stabilized IP weights. 

```{r}
treat_mod <- glm(qsmk ~ sex + age,
                 data = d,
                 family = "binomial")

d$pX <- predict(treat_mod, type = "response")

pn <- glm(qsmk ~ 1, 
          data = d,
          family = "binomial")

d$pnX <- predict(pn, type = "response")
  
d$sw <- with(d, ifelse(qsmk==1, pnX/pX, (1-pnX)/(1-pX)))

```

We can then plot the sample before and after weighting.

```{r, fig.height=3, fig.width=5}
library(ggplot2)
library(patchwork)

p1 <- ggplot() + 
  # X = 1 (sample)
  geom_density(data = subset(d, qsmk == 1), 
                 aes(x = pX), size = 1) +  
  # X = 0 (sample)
  geom_density(data = subset(d, qsmk == 0), 
                 aes(x = pX), linetype = "dashed", size = 1) +
  theme_classic() + 
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.line.y = element_blank()) + 
  xlim(c(0,1)) + xlab("Probability of treatment") +
  ggtitle("Before IP weighting")

p2 <- ggplot() + 
  # X = 1 (pseudo-population)
  geom_density(data = subset(d, qsmk == 1), 
                 aes(x = pX, weight = sw), size = 1) + 
  # X = 0 (pseudo-population)
  geom_density(data = subset(d, qsmk == 0), 
                 aes(x = pX, weight = sw), linetype = "dashed", size = 1) +
  theme_classic() + 
  theme(
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.title.y = element_blank(),
  axis.line.y = element_blank()) + 
  xlim(c(0,1)) + xlab("Probability of treatment") +
  ggtitle("After IP weighting") 

(p1 + p2)
```

We can also make a 'love plot' using the `cobalt` package [@cobalt] to inspect whether the IP weights ensures acceptable balance on the level of individual covariates. By setting `continuous = "std"`, we indicate that the function should return the *standardized* absolute mean difference for any continuous variables (here, `age`). If we wanted the raw absolute mean difference, we'd set `continuous = "raw"`.

```{r, fig.height=3, fig.width=5}
library(cobalt)

love.plot(treat_mod, abs = TRUE,
          sample.names = c("Unweighted", "IP Weighted"), 
          weights = d$sw,
          colors = c("grey60", "black"),
          thresholds = c(m = .1))

bal.tab(treat_mod, abs = TRUE, un = TRUE, thresholds = c(m = .1), weights = d$sw, continuous = "std")$Balance
```

Finally, we include the stabilized weights in an outcome model, which we in turn use for g-computation.

```{r, results='markup'}
out_mod <- lm(wt82_71 ~ qsmk + sex + age, data = d, weights = sw)

EX1 <- predict(out_mod, 
               newdata = transform(d, qsmk = 1))

EX0 <- predict(out_mod, 
               newdata = transform(d, qsmk = 0))

mean(EX1)-mean(EX0)
```

### Bootstrapping

The basic approach to bootstrapping is similar as in the previous chapter. Here, we bootstrap the doubly robust estimator from above. We use only 100 bootstrap samples, but in practice we'd often want more.

```{r, results='markup', cache=TRUE}
library(boot)

# Number of bootstrap samples
n_bootstrap <- 100

bootstrap_analysis <- function(data, indices) {
    
    # Resample the data
    d <- data[indices, ]
  
    # IPW
    treat_mod <- glm(qsmk ~ sex + age,
                     data = d,
                     family = "binomial")

    d$pX <- predict(treat_mod, type = "response")

    pn <- glm(qsmk ~ 1, 
          data = d,
          family = "binomial")

    d$pnX <- predict(pn, type = "response")
  
    d$sw <- with(d, ifelse(qsmk==1, pnX/pX, (1-pnX)/(1-pX)))
    
    # G-computation with IP weighted outcome model
    out_mod <- lm(wt82_71 ~ qsmk + sex + age, data = d, weights = sw)

    EX1 <- predict(out_mod, 
                   newdata = transform(d, qsmk = 1))
  
    EX0 <- predict(out_mod, 
                   newdata = transform(d, qsmk = 0))

    mean(EX1)-mean(EX0)
  
  # Return the coefficient of X
  return(mean(EX1)-mean(EX0))
}

# Perform bootstrapping
bootstrap_results <- boot(data = d, 
                          statistic = bootstrap_analysis, 
                          R = n_bootstrap)

# Summarize the bootstrap results
bootstrap_summary <- boot.ci(bootstrap_results, type = "norm")

# Print the results
print(bootstrap_summary)

```

### More covariates

We can try the same analysis but with a more comprehensive set of covariates.

```{r, results='markup', cache=TRUE}
library(boot)

bootstrap_analysis <- function(data, indices) {
    
    # Resample the data
    d <- data[indices, ]
  
    # IPW
    # see: https://remlapmot.github.io/cibookex-r/ip-weighting-and-marginal-structural-models.html
    treat_mod <- glm(qsmk ~ sex + race + age + I(age ^ 2) + 
                       as.factor(education) + smokeintensity +
                       I(smokeintensity ^ 2) + smokeyrs + I(smokeyrs ^ 2) +
                       as.factor(exercise) + as.factor(active) + wt71 + I(wt71 ^ 2),
                     data = d,
                     family = "binomial")

    d$pX <- predict(treat_mod, type = "response")

    pn <- glm(qsmk ~ 1, 
          data = d,
          family = "binomial")

    d$pnX <- predict(pn, type = "response")
  
    d$sw <- with(d, ifelse(qsmk==1, pnX/pX, (1-pnX)/(1-pX)))
    
    # G-computation with IP weighted outcome model
    out_mod <- lm(wt82_71 ~ qsmk + sex + race + age + I(age ^ 2) + 
                       as.factor(education) + smokeintensity +
                       I(smokeintensity ^ 2) + smokeyrs + I(smokeyrs ^ 2) +
                       as.factor(exercise) + as.factor(active) + wt71 + I(wt71 ^ 2), 
                  data = d, weights = sw)

    EX1 <- predict(out_mod, 
                   newdata = transform(d, qsmk = 1))
  
    EX0 <- predict(out_mod, 
                   newdata = transform(d, qsmk = 0))

    mean(EX1)-mean(EX0)
  
  # Return the coefficient of X
  return(mean(EX1)-mean(EX0))
}

# Perform bootstrapping
bootstrap_results <- boot(data = d, 
                          statistic = bootstrap_analysis, 
                          R = n_bootstrap)

# Summarize the bootstrap results
bootstrap_summary <- boot.ci(bootstrap_results, type = "norm")

# Print the results
print(bootstrap_summary)

```

The overall inference is the same, although the more comprehensive adjustment set yields a slightly higher point estimate (around 3.5 kg), indicating that quitters gain even more weight than previously estimated.

## Bootstrapped sub-group analysis

```{r, results='markup', cache=TRUE}
bootstrap_analysis <- function(data, indices) {
    
    # Resample the data
    d <- data[indices, ]
  
    # IPW
    pn_sub <- glm(qsmk ~ 1 + sex, data = d, family = "binomial")
    
    d$pnX <- predict(pn_sub, type = "response")
  
    d$sw <- with(d, ifelse(qsmk == 1, pnX / pX, (1 - pnX) / (1 - pX)))
    
    # G-computation with IP weighted outcome model
    out_mod <- glm(wt82_71 ~ qsmk + sex + age + qsmk * sex, data = d, weights = sw)
    
    EX1S1 <- predict(out_mod, newdata = transform(d, qsmk = 1, sex = as.factor(1)))
    EX1S0 <- predict(out_mod, newdata = transform(d, qsmk = 1, sex = as.factor(0)))
    EX0S1 <- predict(out_mod, newdata = transform(d, qsmk = 0, sex = as.factor(1)))
    EX0S0 <- predict(out_mod, newdata = transform(d, qsmk = 0, sex = as.factor(0)))
    
    mean_diff_S1 <- mean(EX1S1) - mean(EX0S1)
    mean_diff_S0 <- mean(EX1S0) - mean(EX0S0)
    
    return(c(mean_diff_S1, mean_diff_S0))
}

# Perform bootstrapping
bootstrap_results <- boot(data = d, statistic = bootstrap_analysis, R = n_bootstrap)

# Extract and display results
boot.ci(bootstrap_results, type = "norm", index = 1) # For females
boot.ci(bootstrap_results, type = "norm", index = 2) # For males

```

## Complex longitudinal designs

In the book, we walk through a g-computation approach to a complex longitudinal data context with a time-varying treatment. Here, we show a stabilized IPW approach. The trick is to compute two sets of weights, one for each time point and adjustment set. These weights can then be combined with multiplication in a single marginal structural model that recovers the simulated true effects of 0.

```{r}
set.seed(1747)
n <- 1e4
U <- rnorm(n, 0, 1)
Z_0 <- rbinom(n, 1, plogis(0.5))
X_0 <- rbinom(n, 1, plogis(0.5 + Z_0 * 0.5))
Z_1 <- rbinom(n, 1, plogis(0.5 + X_0 * 0.5 + U * 0.5))
X_1 <- rbinom(n, 1, plogis(0.5 + Z_1 * 0.5))
Y <- rnorm(n, 10 + U * 2)

dat <- data.frame(Y = Y, X_0 = X_0, X_1 = X_1,
                  Z_0 = Z_0, Z_1 = Z_1, U = U)
```

```{r}
# IPW estimation for X_0
# Fit propensity score model (denominator)
ps_model_X0 <- glm(X_0 ~ Z_0 + X_1, family = "binomial", data = dat)
dat$ps_X0 <- predict(ps_model_X0, type = "response")

# Fit numerator model
num_model_X0 <- glm(X_0 ~ Z_0, family = "binomial", data = dat)
dat$num_ps_X0 <- predict(num_model_X0, type = "response")

# Calculate stabilized weights
dat$sw_X0 <- with(dat, 
                ifelse(X_0 == 1, 
                num_ps_X0/ps_X0, 
                (1-num_ps_X0)/(1-ps_X0)))
```

```{r}
# IPW estimation for X_1
# Fit propensity score model (denominator)
ps_model_X1 <- glm(X_1 ~ X_0 + Z_1, family = "binomial", data = dat)
dat$ps_X1 <- predict(ps_model_X1, type = "response")

# Fit numerator model
num_model_X1 <- glm(X_1 ~ X_0, family = "binomial", data = dat)
dat$num_ps_X1 <- predict(num_model_X1, type = "response")

# Calculate stabilized weights
dat$sw_X1 <- with(dat, 
                ifelse(X_1 == 1, 
                num_ps_X1/ps_X1, 
                (1-num_ps_X1)/(1-ps_X1)))
```

```{r, results='markup'}
# Marginal structural model
lm(Y ~ X_0 + X_1, data = dat, weights = sw_X0*sw_X1)
```

## More complexity

In the book, we show a complicated DAG adapted from @vanderweele2016causal of a complex longitudinal exposure-outcome feedback setting. Here, we verify that the adjustment strategy suggested in the book holds true in a simulated setting. While in this particular simulated example the model coefficients for *X1*, *X2* and *X3* in their respective focal models coincide with the marginal estimates we're after, we want to practice a more general workflow for when it really matters. So, we both show a g-computation approach and a stabilized IPW approach.

First, we simulate some data consistent with the complex DAG.

```{r}
# Seed for reproducibility
set.seed(42)

# Define sample size
n <- 1e4

# Simulate time-varying relationships
C <- rnorm(n)
Z1 <- rnorm(n, C) + rnorm(n)
X1 <- rbinom(n, 1, plogis(C + rnorm(n)))
X2 <- rbinom(n, 1, plogis(C + X1 + Z1 + rnorm(n)))
Z2 <- rnorm(n, C + Z1 + X1) + rnorm(n)
X3 <- rbinom(n, 1, plogis(C + X2 + Z2 + rnorm(n)))
Z3 <- rnorm(n, C + Z2 + X2) + rnorm(n)

# Simulate outcome
Y <- X1 + X2 + X3 + Z3 + C + rnorm(n)

# Create dataset
d <- data.frame(C, Z1, Z2, Z3, X1, X2, X3, Y)
```

Next, we fit a model for each measurement time point, and we see that all three models pick up the true effects within simulation error. The true effects are 2, 2 and 1, respectively, for the three time points. It's 2 for the first two time points, because these effects include paths running through *Z2*/*Z3*.

### G-computation

There are no new tricks here compared to what we showcase in the book, except we're breaking down our joint effect estimand into separate models.

```{r, results='markup'}
# Model to estimate effect of X1 on Y
model_X1 <- glm(Y ~ X1 + X2 + Z1 + C, data = d)

EX11 <- predict(model_X1, newdata = transform(d, X1 = 1))

EX10 <- predict(model_X1, newdata = transform(d, X1 = 0))

mean(EX11 - EX10)
```

```{r, results='markup'}
# Model to estimate effect of X2 on Y
model_X2 <- lm(Y ~ X1 + X2 + X3 + Z1 + Z2 + C, data = d)

EX21 <- predict(model_X2, newdata = transform(d, X2 = 1))

EX20 <- predict(model_X2, newdata = transform(d, X2 = 0))

mean(EX21 - EX20)
```

```{r, results='markup'}
# Model to estimate effect of X3 on Y
model_X3 <- lm(Y ~ X2 + X3 + Z2 + C, data = d)

EX31 <- predict(model_X3, newdata = transform(d, X3 = 1))

EX30 <- predict(model_X3, newdata = transform(d, X3 = 0))

mean(EX31 - EX30)
```

### IPW

We follow the same recipe for stabilized IPW as given in the book. 

```{r, results='markup'}
# IPW estimation for X1
# Fit propensity score model (denominator)
ps_model_X1 <- glm(X1 ~ C + X2 + Z1, family = "binomial", data = d)
d$ps_X1 <- predict(ps_model_X1, type = "response")

# Fit numerator model
num_model_X1 <- glm(X1 ~ C, family = "binomial", data = d)
d$num_ps_X1 <- predict(num_model_X1, type = "response")

# Calculate stabilized weights
d$sw_X1 <- with(d, 
                ifelse(X1 == 1, 
                num_ps_X1/ps_X1, 
                (1-num_ps_X1)/(1-ps_X1)))

# Marginal structural model
lm(Y ~ X1 + C, data = d, weights = sw_X1)
```

```{r, results='markup'}
# IPW estimation for X2
# Fit propensity score model (denominator)
ps_model_X2 <- glm(X2 ~ X1 + X3 + Z1 + Z2 + C, family = "binomial", data = d)
d$ps_X2 <- predict(ps_model_X2, type = "response")

# Fit numerator model
num_model_X2 <- glm(X2 ~ C, family = "binomial", data = d)
d$num_ps_X2 <- predict(num_model_X2, type = "response")

# Calculate stabilized weights
d$sw_X2 <- with(d, 
                ifelse(X2 == 1, 
                num_ps_X2/ps_X2, 
                (1-num_ps_X2)/(1-ps_X2)))

# Marginal structural model
lm(Y ~ X2 + C, data = d, weights = sw_X2)
```

```{r, results='markup'}
# IPW estimation for X3
# Fit propensity score model (denominator)
ps_model_X3 <- glm(X3 ~ X2 + Z2 + C, family = "binomial", data = d)
d$ps_X3 <- predict(ps_model_X3, type = "response")

# Fit numerator model
num_model_X3 <- glm(X3 ~ C, family = "binomial", data = d)
d$num_ps_X3 <- predict(num_model_X3, type = "response")

# Calculate stabilized weights
d$sw_X3 <- with(d, 
                ifelse(X3 == 1, 
                num_ps_X3/ps_X3, 
                (1-num_ps_X3)/(1-ps_X3)))

# Marginal structural model
lm(Y ~ X3 + Z3 + C, data = d, weights = sw_X3)
```

Instead of fitting a marginal structural model (MSM) for each time point, we can fit a single MSM for all three time points in one go. The trick here is to multiply the weights. However, this targets a slightly different estimand, since we want to adjust for *Z3* to get a more precise estimate for *X3* but including *Z3* in the MSM blocks the the effect of *X1* and *X2* that runs through *Z3*. That is, this alternative specification targets only the direct effect of the exposures.

```{r, results='markup'}
lm(Y ~ X1 + X2 + X3 + Z3 + C, data = d, weights = sw_X1*sw_X2*sw_X3)
```

For further details on estimating time-varying relationships with IPW, see @vanderweele2016causal.

## Session info

```{r, results='markup'}
sessionInfo()
```

