# Multilevel Modelling and Mundlak's Legacy {#chapter-7}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      results = 'hide', 
                      cache = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      error = TRUE,
                      cache.comments = FALSE,
                      cache.lazy = FALSE,
                      fig.path = "bookdown-dag_files/figure-html/ch7-fig-",
                      cache.path = "bookdown-dag_cache/html/ch7-")

```

```{r}
library(brms)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidybayes)
```


## Multilevel malaria medicine

To reproduce the figure comparing random and fixed effects, we first load the malaria subsidy data and fit the five models. We again use the `brms` package with default prior settings, but models could just as well be implemented using e.g., `lme4`.

```{r}
d <- read.csv("data/cohen2015.csv")
d$stratum <- as.factor(d$stratum)
```

### Naïve model: Simple intercept and slope
```{r, eval=FALSE}
msimp <- brm(act ~ 1 + subsidy,
             data = d, 
             cores = 4, seed = 1,
             file = "fits/msimp")
```

```{r msimp, cache=TRUE, echo=FALSE}
msimp <- readRDS("fits/msimp.rds")
```

### Fixed effects
```{r, eval=FALSE}
mfix <- brm(act ~ 1 + subsidy + stratum,
            data = d,
            cores = 4, seed = 1,
            file = "fits/mfix.rds")
```

```{r mfix, cache=TRUE, echo=FALSE}
mfix <- readRDS("fits/mfix.rds")
```

### Fixed effects interacting treatment and group
```{r, eval=FALSE}
mfix2 <- brm(act ~ 1 + subsidy * stratum,
            data = d,
            cores = 4, seed = 1,
            file = "fits/mfix2.rds")
```

```{r mfix2, cache=TRUE, echo=FALSE}
mfix2 <- readRDS("fits/mfix2.rds")
```

### Random intercepts
```{r, eval=FALSE}
mran <- brm(act ~ 1 + subsidy + (1 | stratum),
            data = d,
            cores = 4, seed = 1,
            file = "fits/mran.rds")
```

```{r mran, cache=TRUE, echo=FALSE}
mran <- readRDS("fits/mran.rds")
```

### Random intercepts and slopes
```{r, eval=FALSE}
mran2 <- brm(act ~ 1 + subsidy + (1 + subsidy | stratum),
            data = d,
            cores = 4, seed = 1,
            file = "fits/mran2.rds")
```

```{r mran2, cache=TRUE, echo=FALSE}
mran2 <- readRDS("fits/mran2.rds")
```

### Plot models in a panel

First, we extract and collect model coefficients in a data frame.

```{r}
FEREpanel <- data.frame(intercept = c(fixef(msimp)[1],
                                      fixef(mran)[1],
                                      coef(mran)$stratum[,1,1],
                                      fixef(mran2)[1],
                                      coef(mran2)$stratum[,1,1],
                                      fixef(mfix)[1],
                                      fixef(mfix)[3:29] + fixef(mfix)[1],
                                      fixef(mfix2)[1],
                                      fixef(mfix2)[3:29] + fixef(mfix2)[1]),
                        slope = c(fixef(msimp)[1],
                                  fixef(mran)[2],
                                  coef(mran)$stratum[,1,2],
                                  fixef(mran2)[2],
                                  coef(mran2)$stratum[,1,2],
                                  rep(fixef(mfix)[2],28),
                                  fixef(mfix2)[2],
                                  fixef(mfix2)[30:56] + fixef(mfix2)[2]),
                        g = c(1, rep(1:29, 2), rep(1:28, 2)),
                        grand = c(1, 1, rep(0,28), 1, rep(0,28), rep(0, 56)),
                        model = c("Y ~ 1 + X",
                                rep("Y ~ 1 + X + (1 | G)", 29),
                                rep("Y ~ 1 + X + (1 + X | G)", 29),
                                rep("Y ~ 1 + X + G", 28),
                                rep("Y ~ 1 + X * G", 28)))
```

Next, we generate predictions from each of the models.

```{r}
# Define a range of x-values (here, it's simply control (X = 0) vs. treatment (X = 1)
x_values <- seq(0, 1, by = 1)

# Expand the dataframe to include x-values for each line...
FEREpanel <- FEREpanel |>
  tidyr::expand_grid(x = x_values) |>
  # ... and compute predictions for control vs. treatment
  dplyr::mutate(y = intercept + slope * x)

```

Then we plot predictions from each of the models

```{r}
# Set y-axis limit
ylim <- c(0,1)

# Generate individual plots
p1 <- FEREpanel |>
  filter(model == "Y ~ 1 + X") |>
  ggplot(aes(x = x, 
             y = y)) +
  geom_line() +
  theme_classic() + 
  labs(subtitle = "Simple intercept and slope",
       title = "Y ~ 1 + X",
       y = NULL, x = NULL) + 
  scale_x_continuous(breaks = c(0,1), 
                     labels = c("Control", "Treatment"), 
                     expand = c(0.1, 0.1)) +
  coord_cartesian(ylim = ylim)

p2 <- FEREpanel |>
  filter(model == "Y ~ 1 + X + G") |>
  ggplot(aes(x = x, 
             y = y, 
             group = g)) +
  geom_line(alpha = 0.15) + 
  theme_classic() + 
  theme(legend.position = "none") +
  labs(subtitle = "Fixed effects",
       title = "Y ~ 1 + X + G",
       y = "Prob. of taking ACT", x = NULL) + 
  scale_x_continuous(breaks = c(0,1), 
                     labels = c("Control", "Treatment"), 
                     expand = c(0.1, 0.1)) +
  coord_cartesian(ylim = ylim)

p3 <- FEREpanel |>
  filter(model == "Y ~ 1 + X * G") |>
  ggplot(aes(x = x, 
             y = y, 
             group = g)) +
  geom_line(alpha = 0.15) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  labs(subtitle = "FE interacting X and group",
       title = "Y ~ 1 + X * G",
       y = NULL, x = NULL) + 
  scale_x_continuous(breaks = c(0,1), 
                     labels = c("Control", "Treatment"), 
                     expand = c(0.1, 0.1)) +
  coord_cartesian(ylim = ylim)

p4 <- FEREpanel |>
  filter(model == "Y ~ 1 + X + (1 | G)") |>
  ggplot(aes(x = x, 
             y = y, 
             group = g, 
             alpha = factor(grand))) +
  geom_line() + 
  scale_alpha_manual(values = c(0.15, 1)) +
  theme_classic() + 
  theme(legend.position = "none") +
  labs(subtitle = "Random intercepts",
       title = "Y ~ 1 + X + (1 | G)",
       y = NULL, x = NULL) + 
  scale_x_continuous(breaks = c(0,1), 
                     labels = c("Control", "Treatment"), 
                     expand = c(0.1, 0.1)) +
  coord_cartesian(ylim = ylim)

p5 <- FEREpanel |>
  filter(model == "Y ~ 1 + X + (1 + X | G)") |>
  ggplot(aes(x = x, 
             y = y, 
             group = g, 
             alpha = factor(grand))) +
  geom_line() + 
  scale_alpha_manual(values = c(0.15, 1)) +
  theme_classic() + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  labs(subtitle = "Random intercepts and slopes",
       title = "Y ~ 1 + X + (1 + X | G)",
       y = NULL, x = NULL) + 
  scale_x_continuous(breaks = c(0,1), 
                     labels = c("Control", "Treatment"), 
                     expand = c(0.1, 0.1)) +
  coord_cartesian(ylim = ylim)
```

Finally, we panel the individual plots using the `patchwork` package

```{r, fig.width=5, fig.height=7}
layout <- "
A#
BC
DE
"

(p1 + p2 + p3 + p4 + p5 + plot_layout(design = layout))
```

### Regularized vs. empirical estimates 

In the text, we show an alternative way to demonstrate partial pooling in a multilevel model. This is where we compare the regularized predictions from a multilevel model against the "empirical" estimates from a fixed effects model. We re-use the model objects from above, `mfix2` and `mran2`, where the treatment effect is allowed to vary by strata.

First, we create two data frames that include the estimated treatment effect for each randomization stratum for each of the two models.

```{r}
# Fixed effects treatment effects
coeffix <- data.frame(
  strata = 1:28,
  coef = fixef(mfix2)[c(2, 30:56), 1])

coeffix$coef[2:28] <- coeffix$coef[2:28] + coeffix$coef[1]

# Random effects treatment effects
coefran <- data.frame(
  strata = 1:28,
  coef = coef(mran2)$stratum[,"Estimate", "subsidy"])

# Compute strata sample sizes
N <- d |>
  group_by(stratum) |>
  summarise(N = n())

# Input strata sample sizes in data frames
coeffix$N <- N$N
coefran$N <- N$N

# Arrange data frames according to strata sample sizes
coeffix <- coeffix |> arrange(N)
coeffix$strata <- factor(coeffix$strata, levels = coeffix$strata)

coefran <- coefran |> arrange(N)
coefran$strata <- factor(coefran$strata, levels = coefran$strata)
```

Then, we plot the stratum-specific treatment effects.

```{r, fig.width=5, fig.height=3}
ggplot() + 
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.07, ymax = 0.3), fill = "lightgrey", alpha = 0.35) +
  geom_hline(yintercept = 0.19, linetype = "dashed", alpha = 0.2, linewidth = 0.7) +
  geom_line(data = rbind(coeffix, coefran), 
            aes(x = strata, y = coef, group = strata),
            linewidth = 0.25,
            alpha = 0.5) +
  geom_point(data = coeffix, aes(x = strata, y = coef),
             size = 1.5,
             alpha = 0.9,
             shape = 1) +  
  geom_point(data = coefran, aes(x = strata, y = coef),
             size = 1.5,
             color = "black") +
  ylab("Treatment Effect") +
  xlab("Randomization Strata\n(smallest to largest)") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank())
```

Instead of using a fixed effects model, an alternative way to compute the "empirical" estimates would be to, within each stratum, take the mean of the outcome variable `act` for the control (`subsidy = 0`) and treatment (`subsidy = 1`) groups separately and then subtract those values.

```{r}
d |>
  # within each stratum and for each subsidy condition (control vs. treatment)...
  group_by(stratum, subsidy) |>
  # ... take the mean of the outcome variable
  summarise("T" = mean(act, na.rm = T)) |>
  # wrangle and name columns
  tidyr::pivot_wider(names_from = "subsidy", 
                     values_from = "T", 
                     names_prefix = "T") |>
  ungroup() |>
  mutate(coef = T1 - T0)
```

## Simulating Mundlak

Here we show the simulation and plotting code for the synthetic Mundlak demonstration.

First, we simulate the confounded data and store it in `d_sim`

```{r}
set.seed(2025)

# define sample size and effects
N_groups <- 30
N_id <- 500
a <- 0
bZY <- 1
bXY <- 0.5

g <- sample(1:N_groups, size = N_id, replace = TRUE) # sample into groups
Ug <- rnorm(N_groups, 1.5, 1) # group confounds
X <- rnorm(N_id, Ug[g], 1) # individual varying trait
Z <- rnorm(N_groups, 0, 1) # group varying trait (observed)
Y <- rnorm(N_id, a + bXY*X + Ug[g] + bZY*Z[g] )

# collect in data frame
d_sim <- data.frame(
  Y = Y,
  X = X,
  Z = Z[g],
  G = as.factor(g)
)
```

To properly quantify uncertainty and obtain neat posterior distributions of the effects from each model, we again analyze the data in a Bayesian framework using `brms` with default priors. We fit our four models.

### Naïve model (ignoring group)
```{r, eval=FALSE}
mNA <- brm(Y ~ X + Z, 
           cores = 4, 
           data = d_sim, seed = 1,
           file = "fits/mNA.rds")
```

```{r mNA, cache=TRUE, echo=FALSE}
mNA <- readRDS("fits/mNA.rds")
```

### Fixed effects
```{r, eval=FALSE}
mFE <- brm(Y ~ X + G, 
           cores = 4, 
           data = d_sim, seed = 1,
           file = "fits/mFE.rds")
```

```{r mFE, cache=TRUE, echo=FALSE}
mFE <- readRDS("fits/mFE.rds")
```

### Random intercepts
```{r, eval=FALSE}
mRE <- brm(Y ~ X + Z + (1 | G), 
           cores = 4, 
           data = d_sim, seed = 1,
           file = "fits/mRE.rds")
```

```{r mRE, cache=TRUE, echo=FALSE}
mRE <- readRDS("fits/mRE.rds")
```

### Mundlak model
```{r, eval=FALSE}
d_sim$Xbar <- with(d_sim, ave(X, G, FUN = mean))

mMU <- brm(Y ~ X + Xbar + Z + (1 | G), 
           cores = 4, 
           data = d_sim, seed = 1,
           file = "fits/mMU.rds")
```

```{r mMU, cache=TRUE, echo=FALSE}
mMU <- readRDS("fits/mMU.rds")
```

### Plot effect estimate distributions

First, we extract posterior distributions of the coefficient of interest (`X`) from each model and collect in a data frame.

```{r}
ndraws <- nrow(brms::as_draws_df(mNA, variable = "b_X"))

forest <- data.frame(bX = c(brms::as_draws_df(mNA, variable = "b_X")$b_X,
                 brms::as_draws_df(mFE, variable = "b_X")$b_X,
                 brms::as_draws_df(mRE, variable = "b_X")$b_X,
                 brms::as_draws_df(mMU, variable = "b_X")$b_X),
           model = c(rep("Naïve model\nY ~ X + Z", ndraws),
                     rep("Fixed effects\nY ~ X + Z + G", ndraws),
                     rep("Random effects\nY ~ X + Z + (1 | G)", ndraws),
                     rep("Mundlak model\nY ~ X + Xbar + Z + (1 | G)", ndraws)))
```

Next, the models are arranged...
```{r}
forest$model <- factor(forest$model,
                       levels=c("Naïve model\nY ~ X + Z",
                                "Fixed effects\nY ~ X + Z + G",
                                "Random effects\nY ~ X + Z + (1 | G)",
                                "Mundlak model\nY ~ X + Xbar + Z + (1 | G)")) |> 
  forcats::fct_rev()
```

... and then plotted.

```{r, fig.width=5, fig.height=3}
forest |>
  ggplot(aes(y = model, x = bX)) +
    stat_halfeye(slab_fill = "white", 
                 slab_color = "grey40", 
                 color = "white") + 
    geom_vline(xintercept = 0.5, 
               linetype = "dashed", 
               alpha = 0.5) + 
    labs(y = NULL, 
         x = "Regression coefficient") +
    scale_x_continuous(breaks = c(0.25, 0.5, 0.75, 1)) +
    theme_classic()
```

## Education and prosociality: Mundlak in action

We finally show the Mundlak model in action in real-world data. We load the `cerc` data [@lang2019moralizing] and plot the raw data distribution for the key predictor and outcome variable, respectively, before fitting and plotting our four models.

```{r}
cerc <- read.csv("data/cerc.csv")
```

### Raw data distributions

```{r, fig.width=5, fig.height=3}
cerc |>
  ggplot(aes(x = FORMALED)) +
  geom_density(aes(y = after_stat(scaled))) +
  facet_wrap(~ SITE, nrow = 2) + 
  labs(y = NULL, x = NULL) +
  scale_x_continuous(n.breaks = 3) +
  scale_y_continuous(breaks = NULL) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  theme_classic() + 
  ggtitle("Years of formal education")
```

```{r, fig.width=5, fig.height=3}
cerc |>
  ggplot(aes(x = Y)) +
  geom_density(aes(y = after_stat(scaled))) +
  facet_wrap(~ SITE, nrow = 2) + 
  labs(y = NULL, x = NULL) +
  scale_x_continuous(n.breaks = 3) +
  scale_y_continuous(breaks = NULL) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  theme_classic() + 
  ggtitle("Coins to co-player")
```

### Naïve model
```{r, eval=FALSE}
mNAcerc <- brm(Y ~ 1 + FORMALED,
               data = cerc,
               cores = 4, seed = 1,
               file = "fits/mNAcerc.rds")
```

```{r mNAcerc, cache=TRUE, echo=FALSE}
mNAcerc <- readRDS("fits/mNAcerc.rds")
```

### Fixed effects
```{r, eval=FALSE}
mFEcerc <- brm(Y ~ 1 + FORMALED+ SITE,
               data = cerc,
               cores = 4, seed = 1,
               file = "fits/mFEcerc.rds")
```

```{r mFEcerc, cache=TRUE, echo=FALSE}
mFEcerc <- readRDS("fits/mFEcerc.rds")
```

### Random effects
```{r, eval=FALSE}
mREcerc <- brm(Y ~ 1 + FORMALED + (1 | SITE),
               data = cerc,
               cores = 4, seed = 1,
               file = "fits/mREcerc.rds")
```

```{r mREcerc, cache=TRUE, echo=FALSE}
mREcerc <- readRDS("fits/mREcerc.rds")
```

### Mundlak model
```{r, eval=FALSE}
cerc$Xbar <- with(cerc, ave(FORMALED, SITE, FUN = mean))

mMUcerc <- brm(Y ~ 1 + FORMALED + Xbar + (1 | SITE),
               data = cerc,
               cores = 4, seed = 1,
               file = "fits/mMUcerc.rds")
```

```{r mMUcerc, cache=TRUE, echo=FALSE}
cerc$Xbar <- with(cerc, ave(FORMALED, SITE, FUN = mean))

mMUcerc <- readRDS("fits/mMUcerc.rds")
```

### Plotting effect of education on religiosity

The code for this plot is much the same as for the simulated Mundlak example above.

```{r}
ndraws <- nrow(brms::as_draws_df(mNAcerc, variable = "b_FORMALED"))

forestcerc <- data.frame(bX = c(brms::as_draws_df(mNAcerc, variable = "b_FORMALED")$b_FORMALED,
                 brms::as_draws_df(mFEcerc, variable = "b_FORMALED")$b_FORMALED,
                 brms::as_draws_df(mREcerc, variable = "b_FORMALED")$b_FORMALED,
                 brms::as_draws_df(mMUcerc, variable = "b_FORMALED")$b_FORMALED),
           model = c(rep("Naïve model", ndraws),
                     rep("Fixed effects", ndraws),
                     rep("Random effects", ndraws),
                     rep("Mundlak model", ndraws)))
  
forestcerc$model <- factor(forestcerc$model,
                           levels=c("Naïve model",
                                    "Fixed effects",
                                    "Random effects",
                                    "Mundlak model")) |> forcats::fct_rev()
```

```{r, fig.width=5, fig.height=3}
forestcerc |>
  ggplot(aes(y = model, 
             x = bX)) +
    stat_halfeye(point_interval = "mean_hdci", 
                 .width = 0.95, 
                 slab_fill = "white", 
                 slab_color = "grey40", 
                 color = "white") + 
    geom_vline(xintercept = 0, 
               linetype = "dashed", 
               alpha = 0.5) + 
    labs(title = "Cross-cultural Dictator Game",
         y = NULL, 
         x = "Effect of Education on Prosociality") +
    theme_classic()

```

## Marginal effects in a multilevel model

As we discuss in the text, there are several ways of obtaining predictions from a multilevel model, including predictions for *each site*, for an *average site* and *marginal of site*. 

Note that we here show a general g-computation approach, where we first compute marginal effects *within* each MCMC draw (`group_by(.draw, ...) |> summarise(.epred = mean(.epred))`), before summarizing by the posterior mean and quantile intervals *across* MCMC draws (`group_by(FORMALED) |> summarise(.epred = mean_qi(.epred))`). In the current data example, the `group_by(.draw, ...)` step is redundant, but it's how we would go about applying g-computation to obtain marginal effects in a setting with a *continuous* focal (causal) predictor and with possible covariates and nonlinearities.

First, we fit an extended Mundlak model, that allows the effect of education on religiosity to vary by site.

```{r, eval=FALSE}
mMUcerc2 <- brm(Y ~ 1 + FORMALED + Xbar + (1 + FORMALED | SITE),
                data = cerc,
                cores = 4, seed = 1,
                file = "fits/mMUcerc2.rds")
```

```{r mMUcerc2, cache=TRUE, echo=FALSE}
mMUcerc2 <- readRDS("fits/mMUcerc2.rds")
```

### Predicting the observed sites

We can then visualize effect estimates for each group (field site, in this case) across the full range of observed education years and at site-specific average years of education.

```{r}
# prepare prediction grid across full range of education years...
nd <- tidyr::expand_grid(FORMALED = c(0,10,20,30),
                         SITE = unique(cerc$SITE))

# ... and at site-specific average years of education
nd$Xbar <- rep(unique(cerc$Xbar), length(unique(nd$FORMALED)))
```

Plot predictions for each site in separate panels. For this, we need to include the random effects in the predictions.

```{r, fig.width=5, fig.height=3}
# Predict effect estimates...
add_epred_draws(mMUcerc2,
                # for prediction grid and...
                newdata = nd,
                # ... *include* all random effect components.
                re_formula = NULL) |>
  
  # Compute average effect *within* each MCMC draw.
  group_by(.draw, SITE, FORMALED) |>
  summarise(.epred = mean(.epred)) |>

  # Summarise average effect *across* each MCMC draw
  # for each site and educational level.
  group_by(SITE, FORMALED) |>
  summarise(mean_qi(.epred)) |>

  # Plot!
  ggplot(aes(x = FORMALED, y = y, ymin = ymin, ymax = ymax)) +
  geom_lineribbon(color = "blue",
                  fill = "grey90",
                  linewidth = 0.5) +
  coord_cartesian(ylim = c(0,8)) +
  facet_wrap(~SITE, nrow = 2) +
  theme_classic() + 
  theme(legend.position = "none") +
  labs(title = "Predicting the observed sites", 
       subtitle = "Including random effects",
       x = "Years of education", y = "Coins to co-player")
```

But we can also ignore site-specific effects altogether and instead aim at predicting a perfectly average site. This amounts to ignoring the random effects of the model. We first need to set up a new prediction grid for the new hypothetical site -- let's call it "Newland" -- with an average of 7 years of education, the average of the site averages (this can be checked by running `mean(unique(cerc$Xbar))`).

```{r}
nd2 <- tidyr::expand_grid(FORMALED = c(0,10,20,30),
                          SITE = "Newland", # could also just set to NA
                          Xbar = 7) # could also make a distribution of Xbars to average over
```

```{r}
# Predict effect estimates...
p1 <- add_epred_draws(mMUcerc2,
                # for new prediction grid and...
                newdata = nd2,
                # ... *ignore* all random effect components.
                re_formula = NA) |>

  # Compute average effect *within* each MCMC draw.
  group_by(.draw, FORMALED) |>
  summarise(.epred = mean(.epred)) |>
  
  # Summarise average effect *across* MCMC draws
  # for each educational level.
  group_by(FORMALED) |>
  summarise(mean_qi(.epred)) |>

  # Plot!
  ggplot(aes(x = FORMALED, y = y, ymin = ymin, ymax = ymax)) +
  geom_lineribbon(color = "blue",
                  fill = "grey90",
                  linewidth = 0.5) +
  coord_cartesian(ylim = c(0,8)) +
  scale_fill_brewer() + 
  theme_classic() + 
  theme(legend.position = "none") +
  labs(title = "Predicting the average site", 
       subtitle = "Ignoring random effects",
       x = "Years of education", y = "Coins to co-player")
```

And finally, we could be interested in predicting a new site drawing from all that we know about the observed sites. This amounts to *averaging over* -- instead of *ignoring*, as above -- the uncertainty that we have around the observed sites and generating predictions from that.

```{r}
# Predict effect estimates...
p2 <- add_epred_draws(mMUcerc2,
                      # for prediction grid and...
                      newdata = nd2,
                      # ... *include* all random effect components.
                      re_formula = NULL,
                      # Allow predictions for an unobserved group and...
                      allow_new_levels = TRUE,
                      # ... sample from the implied multivariate gaussian.
                      sample_new_levels = "gaussian") |>
  
  # Compute average effect *within* each MCMC draw.
  group_by(.draw, FORMALED) |>
  summarise(.epred = mean(.epred)) |>
  
  # Summarise average causal effect across MCMC draws
  # for each educational level.
  group_by(FORMALED) |>
  summarise(mean_qi(.epred)) |>

  # Plot!
  ggplot(aes(x = FORMALED, y = y, ymin = ymin, ymax = ymax)) +
  geom_lineribbon(color = "blue",
                  fill = "grey90",
                  linewidth = 0.5) +
  coord_cartesian(ylim = c(0,8)) +
  scale_fill_brewer() + 
  theme_classic() + 
  theme(legend.position = "none") +
  labs(title = "Predicting a new site", 
       subtitle = "Averaging over random effects",
       x = "Years of education", y = NULL) + 
  scale_y_continuous(breaks = NULL)
```

```{r, fig.width=5, fig.height=3}
p1 + p2 + plot_layout(axis_titles = "collect")
```

### Frequentist workflow

For completeness, we can also plot predictions for each site with a frequentist approach.

First, we fit a corresponding frequentist model. 

```{r}
library(lme4)

mMUcerc2_freq <- lmer(Y ~ 1 + FORMALED + Xbar + (1 + FORMALED | SITE),
                      data = cerc)
```

Next, we obtain predictions using the `marginaleffects` package to obtain confidence intervals. Now, the frequentist approach excludes uncertainty in the random effects, so it's not entirely comparable to the predictions obtained in the text.

```{r, fig.width=5, fig.height=3}
library(marginaleffects)

# Predicting the observed sites (only with uncertainty in the global/fixed effects parameters)
predictions(mMUcerc2_freq,
            newdata = nd) |>
  ggplot(aes(x = FORMALED, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_lineribbon(color = "blue",
                  fill = "grey90",
                  linewidth = 0.5) +
  coord_cartesian(ylim = c(0,8)) +
  facet_wrap(~SITE, nrow = 2) +
  theme_classic() + 
  theme(legend.position = "none") +
  labs(title = "Predicting the observed sites", 
       subtitle = "Frequentist version",
       x = "Years of education", y = "Coins to co-player")

```

## Session info

```{r, results='markup'}
sessionInfo()
```
